Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-21 inference.tex
DSL syntax made same as OCaml syntax.
add_interest example better explained.

states, respectively.  The language is carefully chosen to be

@@

transformer $\F'$ that simply ignores its $\stg$ argument in favor of a generic
$\stg'$, making $\F'$ trivially stable. It is safe to assume

@@

\rulelabel{RG-Conseq} rule for transaction-bound commands.  The weakening semantics chosen by
$\stabilize{\cdot}$, while being simple, is nonetheless useful because
of the $I(\stg')$ assumption on the existentially bound $\stg'$. The
example in Fig.~\ref{fig:weakening-example} demonstrates. 

@@

let add_interest acc_id pc = atomically @@ do
  let a = SQL.select1 BankAccount (fun acc -> acc.id = acc_id);
  let y = a.bal + pc*a.bal;

@@

(\C{pc}) to the balance of a bank account, which is required to be
non-negative ($I(\stg) \Leftrightarrow \forall(r\in\stg).~r.\C{bal}\ge
0$). The transaction starts by issuing a \C{select1} query, whose
effect implicitly asserts that there exists a record in the bank
account database that is equal to $a$ ($\exists(r\in\stg).~r=a$).
However, this assertion is unstable because a concurrent \C{withdraw}
or \C{deposit} transactions might update the account balance, so that
such a record no longer exists in the database.  Fortunately, the
weakening operator ($\stabilize{\cdot}$) allows us to weaken the
assertion to $\exists \stg'.I(\stg') \conj \exists(r\in\stg').~a=r$
({\sf exists} has the same semantics as $\exists$, as formalized in
Fig.~\ref{fig:logic}), which is enough to prove that $\C{a.bal +
pc*a.bal}\ge 0$, and verify that \C{add\_interest} preserves the
positive balance invariant.

>>>>

states, respectively. Constant sets can be written using the set
literatl expressions. For e.g., the set $\{1,2\}$ can be written as
$\{x \,|\, x=1 \vee x=2\}$. The language is carefully chosen to be

@@

transformer $\F'$ that simply ignores its $\stg$ argument in favor of
a generic $\stg'$, making $\F'$ trivially stable. It is safe to assume

@@

\rulelabel{RG-Conseq} rule for transaction-bound commands.  The
weakening semantics chosen by $\stabilize{\cdot}$, while being simple,
is nonetheless useful because of the $I(\stg')$ assumption on the
existentially bound $\stg'$. The example in
Fig.~\ref{fig:weakening-example} demonstrates. 

@@

let add_interest acc_id pc = atomically_do @@ fun () ->
  let a = SQL.select1 BankAccount (fun acc -> acc.id = acc_id) in
  let y = a.bal + pc*a.bal in

@@

(determined by \C{pc}) to the balance of a bank account, which is
required to be non-negative ($I(\stg) \Leftrightarrow
\forall(r\in\stg).~r.\C{bal}\ge 0$). The transaction starts by issuing
a \C{select1} query, whose effect ($\F$) is essentially a singleton
set containing a record $r$ whose id is \C{acc\_id} (i.e.,
$\F(\stl,\stg) = \{r \,|\, r\in\stg \wedge r.\C{id} = \C{acc\_id}\}$).
However, this effect is unstable because $\F(\stl,\stg')$ may not be
the same set as $\F(\stl,\stg)$ when $\stg'\neq\stg$. A record
$r\in\stg$ whose $\C{id}=\C{acc\_id}$ may have its balance updated by
a concurrent \C{withdraw} or \C{deposit} transaction in $\stg'$,
making the record in $\stg'$ different from the record in $\stg$.
Hence the stability check fails.  Fortunately, the weakening operator
($\stabilize{\cdot}$) allows us to weaken the effect to
$\existsl(\stg, I(\stg), \{ r \,|\, r\in\stg \wedge
r.\C{id}=\C{acc\_id}\})$, which effectively asserts that the
\C{select1} query returns a record with $\C{id}=\C{acc\_id}$ from
\emph{some} database state that satisfies the non-negative balance
invariant $I$.  This weakened assertion is nonetheless enough to
deduce that $\C{a.bal}\ge0$, and subsequently prove that $\C{a.bal +
pc*a.bal}\ge 0$, allowing us to verify the \C{add\_interest}
transaction.
